---
title: "Task 2"
author: "Max Settineri"
date: "2023-01-31"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(AICcmodavg)
library(tidyverse)
library(here)
library(equatiomatic)
library(kableExtra)
```

```{r}
calcofi <- read_csv(here('data', 'calcofi_seawater_samples.csv')) 
```

## Multiple Linear Regression Models

Creating two multiple linear regression models:

- Oxygen saturation as a function of water temperature, salinity, and phosphate concentration
- Oxygen saturation as a function of water temperature, salinity, phosphate concentration, and depth

```{r}
f1 <- o2sat ~ t_deg_c + salinity + po4u_m
model1 <- lm(f1, data = calcofi)

f2 <- o2sat ~ t_deg_c + salinity + po4u_m + depth_m
model2 <- lm(f2, data = calcofi)
```

## Using AIC to identify the better model

```{r}
aictab(list(model1, model2))
```

## Using BIC to identify the better model

```{r}
bictab(list(model1, model2))
```

## Comparing models using ten-fold cross validation

```{r}

```

