---
title: "Task 1"
author: "Max Settineri"
date: "2023-01-26"
output: html_document
---
### Introduction
Brief overview

**Data citation:** Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3

### Setup 
```{r setup, include = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(lubridate)


```

```{r}
# reading in data and using lubridate to add a new column for year
amphibians <- read_excel(here('data', 'sierra_amphibians.xlsx')) %>% 
  mutate(year = year(survey_date))
```

### Visualizing Mountain Yellow-Legged Frog Counts by Year and Life Stage

```{r}
frog <- amphibians %>% 
  filter(amphibian_species == 'RAMU') %>% # including only RAMU species
  filter(amphibian_life_stage %in% c('Adult', 'SubAdult', 'Tadpole')) %>% # excluding the life stage EggMass
  group_by(amphibian_life_stage, year) %>% 
  summarize(total_count = n()) # getting counts for the number of species in each life stage by year

```

```{r}
# plotting the total RAMU counts by year and life stage
graph1 <- ggplot(data = frog, aes(x = year, y = total_count, fill = amphibian_life_stage)) +
  geom_col() +
  labs(x = 'Year', y = 'Yellow-Legged Frog Count') +
  scale_fill_manual(values = c('darkgreen', 'seagreen3', 'palegreen')) +
  theme_minimal()

graph1
```

### Visualizing Mountain Yellow-Legged Frog Counts by Lake

```{r}
lake <- amphibians %>% 
  filter(amphibian_species == 'RAMU') %>% # including only RAMU species
  filter(amphibian_life_stage %in% c('Adult', 'SubAdult')) %>% # including only adults/subadult
  group_by(lake_id) %>% # grouping by lake
  summarize(total_count = n()) %>% # getting total counts of observations at each lake
  mutate(lake_id = paste0("Lake ", lake_id)) %>% # adding 'Lake' text to lake_id's
  mutate(lake_id = as.factor(lake_id)) %>% # converting lake IDs to factor data
  slice_max(order_by = total_count, n = 5) %>% # including only the 5 lakes with the most observations
  arrange(desc(total_count))
```

```{r}
# making a column graph with the total counts from the 5 lakes with the most observations
graph2 <- ggplot(data = lake, aes(x = fct_reorder(lake_id, total_count, .desc = TRUE), y = total_count)) +
  geom_col(aes(fill = lake_id)) +
  labs(x = ' ', y = 'Yellow-Legged Frog Count') +
  theme_minimal(10) +
  scale_fill_manual(values = c('steelblue2', 'skyblue1', 'steelblue4', 'slategray1', 'steelblue3')) +
  theme(legend.position = "none")

graph2
```

